---
title: "homework_03"
author: "Sanjana Sujeet"
date: "2025-05-28"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(janitor)
library(here)
library(gt)
library(dplyr)
library(ggplot2)
library(tinytex)
```


**GitHub Repository:** [https://github.com/sanjanasujeet/ENVS-193DS_workshop-07](https://github.com/sanjanasujeet/ENVS-193DS_workshop-07)

# 1. Personal Data
## 1A. Data Summarizing

My data tracked hours I read, pages I read, and social hours per day as my response variables and stress level as my predictor variable; the main thing that I want to look at is how stress levels affect my reading.  

I will calculate my reading pace by dividing pages read by hours reading and compare this to my quantified daily stress level, which I calculated based on the number of upcoming assignments and tests as well as my perceived stress, to see if stress influences the pace that I read.

To summarize my data, I could calculate the reading pace across different stress levels (low, medium, high stress) to see if stress impacts not how much I read, but how fast I read. Comparing means within stress levels is informative because it is a simple way to see if there is a difference between how different stress levels affect me. I may read less efficiently when I feel overwhelmed, or I may quicker when trying to avoid stress through a calming activity.

#### ignore-cleaning data

```{r}
data<- read.csv("data/springtracking2.csv") #read in data
data_clean<-clean_names(data) #cleaning column names
data_clean <- data_clean |> 
  rename(hours = hours_of_activity) #shortens name cuz i think i will be using it often
data_clean <- data_clean |> 
  mutate(pace = pages_read / hours) #creates new column pace which is pages/min
data_clean <- data_clean %>%
  mutate(pace = replace(pace, is.nan(pace), 0))# removes NaN and makes them 0 instead(this was a calculation issue)
data_clean <- data_clean %>%
  mutate(stress_level = case_when(
    quantified_stress >= 0 & quantified_stress <= 9  ~ "low",
    quantified_stress >= 10 & quantified_stress <= 19 ~ "medium",
    quantified_stress >= 20 ~ "high")) # creates new column called stress_level and takes quantified stress levels and makes 3 levels out of it

data_clean <- data_clean %>%
  mutate(stress_level = factor(stress_level, levels = c("low", "medium", "high"))) # orders the levels

```

#### ignore-preliminary graphs

```{r, message=FALSE, echo=FALSE, eval= FALSE}
#graphing

ggplot(data_clean %>% 
         filter(!is.na(quantified_stress), !is.na(pages_read)),
       aes(x = quantified_stress, y = pages_read)) +
  geom_point(color = "cyan4", alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", color = "royalblue", se = TRUE) +
  labs(
    title = "Pages Read vs. Quantified Stress",
    x = "Quantified Stress",
    y = "Pages Read"
  ) +
  theme_minimal()

```

```{r, message=FALSE, echo=FALSE, eval= FALSE}
ggplot(data_clean |> filter(!is.na(stress_level), !is.na(pages_read)),
       aes(x = stress_level, y = pages_read, fill = stress_level)) +
  geom_boxplot() +
  scale_fill_manual(values = c("low" = "lightblue", "medium" = "cadetblue", "high" = "cyan4")) +
  labs(
    title = "Pages Read by Stress Level",
    x = "Stress Level",
    y = "Pages Read"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

```{r, message=FALSE}
ggplot(data_clean %>% 
         filter(!is.na(quantified_stress), !is.na(pace), pace != 0),
       aes(x = quantified_stress, y = pages_read)) +
  geom_point(color = "salmon", alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", color = "darkred", se = TRUE) +
  labs(
    title = "Pace vs. Quantified Stress (Excluding no reading days)",
    x = "Quantified Stress",
    y = "Pace(pages per hour)"
  ) +
  theme_minimal()

```

## 1B. Visualization

```{r}
ggplot(data_clean |> filter(!is.na(stress_level), !is.na(pace), pace != 0),
       aes(x = stress_level, y = pace, fill = stress_level)) +
  geom_boxplot() +
  scale_fill_manual(values = c("low" = "rosybrown1", "medium" = "salmon", "high" = "indianred4")) +
  labs(
    title = "Reading Pace by Stress Level (Excluding no reading days)",
    x = "Stress Level",
    y = "Pace(Pages per Hour)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

## 1C. Caption
**Figure: Reading Pace by Stress Level (Excluding Days with No Reading)**

This boxplot shows the distribution of reading pace, measured in pages per hour, across three calculated stress levels: low, medium, and high. During days of low stress levels there is high variability and highest median pace; days of high stress consistently read fewer pages per hour than medium stress level days with a more compressed range. The trend suggests that increased stress might negatively affect reading pace and consistency but the large range of pace during low stress days .

## 1D. Table Presentation
```{r}
# Summarize stats of reading pace by stress level
summary_table <- data_clean %>%
  filter(pace != 0) %>%
  group_by(stress_level) %>%
  summarise(
    Min = round(min(pages_read), 1),
    Q1 = round(quantile(pace, 0.25), 1),
    Median = round(median(pace), 1),
    Q3 = round(quantile(pace, 0.75), 1),
    Max = round(max(pages_read), 1)  ) %>%
  arrange(factor(stress_level, levels = c("low", "medium", "high")))
#create gt table
summary_table %>%
  gt() %>%
  tab_header(title = "Distribution of Reading Pace by Stress Level") %>%
  cols_label(
    stress_level = "Stress Level",
    Min= "Minimum",
    Q1 = "1st Quartile",
    Median = "Median",
    Q3 = "3rd Quartile",
    Max= "Maximum")

```

# 2. Affective Visualization
## 2A. Describe in words what an affective visualization could look like for your personal data

I could create a calendar with books instead of regular squares and they are colored in by stress level with a number of the pace on each book. I can add in notes in little bubbles connected to the books like bookmarks for the days I wrote observations like 're-read the same page 3 times' or 'sped through this.'

## 2B. Sketch

## 2C. Visualization Draft

## 2D. Artist Statement

This piece visualizes my personal reading pace across days of with stress levels. Each book icon represents a day I read, color-coded by stress level(green-low, blue- medium, coral- high) and labeled with how many pages I read per hour.

I was inspired by my life's saviour: google calender, and bullet journal youtube creators like Erin Smith and KB Journals. I track everything in my life down to the minute with my google calendar and I love how easy it is to just look at the page and know a lot of information through color coding, time blocks, and writing. But before I used gcal, I used bullet journals and the idea of using books as calendar units came from an old bullet journal spread I had in middle school, where I tracked which books I read which days(back when I read multiple books in a week.)

The final piece is a digital hand-drawn calendar with color codes and written info made through Procreate.

I started with a template of books in a grid like a calendar and use the google sheet I made to add colors and pace to each book. Then using the Notes column I input short sentences about my qualitative observations and tried to make them look like bookmarks. 










